---
title: "Testing Sankay Charts and other Graphs of Graphs and Trees"
output: html_notebook
---



```{r}
library(tidyverse)
library(networkD3)
```

# Testing The Sankey Charts from `networkD3`
The following example is modified from a stack overflow post:
https://stackoverflow.com/questions/52895046/structuring-binary-data-for-sankey-plot/52895853#52895853

## Generate Random Data
```{r}
set.seed(900)
n=100
example.data<-data.frame(A = rep(1,n),
                         B = sample(c(0,1),n,replace = TRUE)) %>%
  mutate(C = sample(c(0,1),n,prob = c(0.3,0.7),replace = TRUE),
         D = sample(c(0,1),n,prob = c(0.85,0.15),replace = TRUE))
```

## Pivot data to long form
```{r}
events <- example.data %>%
  mutate(row = row_number()) %>% 
  pivot_longer(-row, names_to = "column", values_to = "choice" ) %>%

  mutate(column_num = match(column, names(example.data))) %>% 
  arrange(row, column_num) %>% 
  select(row, column_num, everything())
```

## Create "target" and "source" columns
The coding is based on the knowledge that if you group by row, the first
item in each row is from column A, and the last was from column D. When you 
you `lag()` the data within that row, the first item will always be `NA`.
We end up with "source" values that have `NA` wherever a column A value would
have been. We then filter out any NAs. I dont believe it is essential to 
filter out the `NA`s....

What I find weird here is the terminology.  I want the "source" to be to the 
left, but that's not the way this package is constructed. The on-line help 
provides no real help in figuring out why.

```{r}
links1 <-
  events %>% 
  mutate(target = paste0(column, "_", choice)) %>% 
  group_by(row) %>% 
  mutate(source = lag(target)) %>%
  ungroup()

links1.1 <- links1 %>% 
  filter(!is.na(source) & !is.na(target))
```

```{r}
links2 <- 
  links1.1 %>% 
  select(source, target) %>% 
  group_by(source, target) %>% 
  summarise(value = n(), .groups = 'drop') %>% 
  ungroup()
```

## Create a Dateframe of Node Names
At this point we have a Links dataframe, but with text names for each node. The
`networkD3` package wants nodes identified numerically, with names isolated in
a separate dataframe. 

I create a dataframe containing names for all the nodes...

```{r}
nodes <- data.frame(name = unique(c(links2$source, links2$target)))
```


## Generate the Plot
Finally, I replace all the node string values in links with numeric codes, zero
indexed.

```{r}
links3 <- links2 %>%
  mutate(source = match(source, nodes$name) - 1,
         target = match(target, nodes$name) - 1)

plt <- sankeyNetwork(Links = links3,
             Source = "source", Target = "target", Value = "value",
             Nodes = nodes, 
             NodeID = "name", nodeWidth = 25, nodePadding = 20,
             fontSize = 48)
plt
```

## Save the Output
The graphic is not easy to save programmatically.  The save functionality
produces an HTML widget.  A problem in  `rmarkdown` functions that are actually
called to generate the saved HTML prevents saving this as a self-contained HTML 
file.

```{r}
saveNetwork(plt, file = 'sankay.html', selfcontained = TRUE)
```

# Testing a Simple Graph from `networkD3`
`simpleNetwork` Generates a force directed network graph, again as an HTML 
Widget.  The following example is from the related help page.
```{r}
src <- c("A", "A", "A", "A", "B", "B", "C", "C", "D")
target <- c("B", "C", "D", "J", "E", "F", "G", "H", "I")
networkData <- data.frame(src, target)

sn <- simpleNetwork(networkData, fontSize = 12)
sn
```

# Testing a Hierarchical Tree from `networkD3`
The following example is from the help file for `diagonalNetwork()`.

`diagonalNetwork()` generates a "Reingold-Tilford Tree" network diagram
```{r}
CanadaPC <- list(name = "Canada", children = list(list(name = "Newfoundland",
                    children = list(list(name = "St. John's"))),
               list(name = "PEI",
                    children = list(list(name = "Charlottetown"))),
               list(name = "Nova Scotia",
                    children = list(list(name = "Halifax"))),
               list(name = "New Brunswick",
                    children = list(list(name = "Fredericton"))),
               list(name = "Quebec",
                    children = list(list(name = "Montreal"),
                                    list(name = "Quebec City"))),
               list(name = "Ontario",
                    children = list(list(name = "Toronto"),
                                    list(name = "Ottawa"))),
               list(name = "Manitoba",
                    children = list(list(name = "Winnipeg"))),
               list(name = "Saskatchewan",
                    children = list(list(name = "Regina"))),
               list(name = "Nunavuet",
                    children = list(list(name = "Iqaluit"))),
               list(name = "NWT",
                    children = list(list(name = "Yellowknife"))),
               list(name = "Alberta",
                    children = list(list(name = "Edmonton"))),
               list(name = "British Columbia",
                    children = list(list(name = "Victoria"),
                                    list(name = "Vancouver"))),
               list(name = "Yukon",
                    children = list(list(name = "Whitehorse")))
))

diagonalNetwork(List = CanadaPC, fontSize = 10)

```


# Testing the `data.tree` Package
This package handles creation of hierarchical data.

Much of the following is from the package vignettes.
```{r}
library('data.tree')
```

```{r}
acme <- Node$new("Acme Inc.")
  accounting <- acme$AddChild("Accounting")
    software <- accounting$AddChild("New Software")
    standards <- accounting$AddChild("New Accounting Standards")
  research <- acme$AddChild("Research")
    newProductLine <- research$AddChild("New Product Line")
    newLabs <- research$AddChild("New Labs")
  it <- acme$AddChild("IT")
    outsource <- it$AddChild("Outsource")
    agile <- it$AddChild("Go agile")
    goToR <- it$AddChild("Switch to R")

print(acme)
```

Plotting is managed through `DiagrammeR`, which has lots of formatting options.
`DiagrammeR` is itself builds on `graphviz`

```{r}
library(DiagrammeR)
plot(acme)
```

## Loading Data from the `treemap` package
```{r}
library(treemap)
data(GNI2014)
head(GNI2014)
```

## Building a Tree based on that Tabular Data
```{r}
GNI2014$pathString <- paste("world", 
                            GNI2014$continent, 
                            GNI2014$country, 
                            sep = "/")
```


The following call o `as.Node()` implicitly calls `FromDataFrameTable()`, where
a lot of the heavy lifting occurs.
```{r}
population <- as.Node(GNI2014)
print(population, "iso3", "population", "GNI", limit = 20)
```


I also read about the `igraph` package....  `igraph` manages graphs, as in graph 
theory -- collections of nodes and links. It's a big package. I've just 
scratched the surface.

```{r}
library(igraph)
ig <- as.igraph(acme, directed = TRUE, direction = "climb")
plot(ig)
```
That is plotted with standard R graphics, so export requires opening a graphics 
device....

```{r}
print(ig)
```



```{r}
g <- make_ring(10, directed = TRUE, mutual = TRUE)
plot(g, layout=layout_with_kk, vertex.color="green")
```
```{r}
l <- make_lattice(c(3, 3, 2))
plot(l)
```

